New insights into the Weddell Sea ecosystem applying a network approach

### 2022-08-10 Manuscript/Manuscript.Rmd
Results: Tables 1 & 2

With the aim of analysing the effect of each species on the food web's stability we deleted one species at a time, so the network size was reduced by one. After each species extinction, we calculated the stability for the food web minus one species (size = 489) and compared it with that of the whole network (size = 490); we performed 1000 simulations for each species extinction. At last we statistically analysed such difference with an Anderson-Darling test considering a p-value < 0.01 [@Scholz1987]. If the difference was positive, then the stability of the food web was impacted positively (higher stability) when that species become extinct, and viceversa. Details for the stability calculations are described in Supplementary Material.

@@

First of all, it's important to note that since the proportion of Jacobians that are locally stable is zero for the Weddell Sea food web, we considered the real part of the maximum eigenvalue as the stability index. Our results showed that the majority of the species had no significant impact on the stability of the food web when extinct (Anderson-Darling test p-value > 0.01 for QSS difference). This is clearly observed in Figure 6 A-D, where most of the points lie around the zero value of the 'QSS difference' axis. This means that the stability for the food web minus a given species is similar to the stability for the whole network.

When plotting

Biplots btw QSS difference and weighted and unweighted prop.

@@

```{=tex}
\begin{table}[t]
\caption{Model comparison for the distribution of interaction strengths of the Weddell Sea food web. According to AIC, log-Normal is the best model.}
\begin{tabular}{l c c c}
\tophline

Model & df & AIC & deltaAIC \\
\middlehline
log-Normal & 2 & -356579 & 0 \\
\middlehline
Gamma & 2 & -352575 & 4004 \\
\middlehline
Power-law & 2 & -347646 & 8933 \\
\middlehline
Exponential & 1 & -262852 & 93726 \\
\middlehline
Normal & 2 & -222785 & 133793 \\
\middlehline
Uniform & 2 & -178001 & 178578 \\

>>>>

With the aim of analysing the effect of each species on the food web's stability we deleted one species at a time, so the network size was reduced by one. After each species extinction, we calculated the stability for the food web minus one species (size = 489) and compared it with that of the whole network (size = 490). We performed 1000 simulations for each species extinction and obtained a mean QSS and maximum eigenvalue. At last we statistically analysed such difference with an Anderson-Darling test considering a p-value < 0.01 [@Scholz1987]. If the difference was positive, then the stability of the food web was impacted positively (higher stability) when that species become extinct, and viceversa. Details for the stability calculations are described in Supplementary Material.

@@

It's important to note that since the proportion of Jacobians that are locally stable or QSS was zero for the Weddell Sea food web, we considered the mean maximum eigenvalue as the stability index. Our results showed that the majority of the species had no significant impact on the stability of the food web when extinct (Anderson-Darling test p-value > 0.01). This is clearly observed in Figure 6 A-D, where most of the points lie around the zero value of the 'Stability difference' axis. This means that the stability for the food web minus a given species is similar to the stability for the whole network. However, there were few species that when become extinct generated positive significant differences in the stability of the food web, meaning that they increased the network's stability when deleted. There is an exception, Pagetopsis macropterus (demersal fish), that when deleted decreased the stability. In total, 13 out of 490 species (2.65%) gave rise to significant changes in the network's stability.

When plotting the stability difference against species properties and the previous results regarding clustering, interesting insights arose (Figure 6). In the first place, all species with a significant impact on the stability belong to the 'High IS' group, denoting a high mean interaction strength (Figure 6A). Secondly, these species are positioned in high trophic levels (TL > 3.7) (Figure 6B). Thirdly, they have a relatively high number of interactions or degree (Degree > 25) (Figure 6C). Regarding trophic similarity, species have mid to low values (TS < 0.13) (Figure 6D). Habitat wise, species live in all the habitats, except for are benthic (Figure 6E). Table 2 shows these results per species.

@@

```{=tex}
\begin{table}[t]
\caption{Model comparison for the distribution of interaction strengths of the Weddell Sea food web. Order by best fit. References: df = degrees of freedom, AIC = Akaike Information Criterion, deltaAIC = difference with best fit. Log-Normal is the best model.}
\begin{tabular}{l c c c}
\tophline

\textbf{Model} & \textbf{df} & \textbf{AIC} & \textbf{deltaAIC} \\
\middlehline
log-Normal & 2 & -356579 & 0 \\
\middlehline
Gamma & 2 & -352575 & 4004 \\
\middlehline
Power-law & 2 & -347646 & 8933 \\
\middlehline
Exponential & 1 & -262852 & 93726 \\
\middlehline
Normal & 2 & -222785 & 133793 \\
\middlehline
Uniform & 2 & -178001 & 178578 \\

```{=tex}
\begin{table}[t]
\caption{Properties of the species that when become extinct generated a significant impact on the stability of the Weddell Sea food web, ordered by significance (Anderson-Darling p value). All species belong to the 'High IS' group.  References: meanIS = mean interaction strength, TL = trophic level, Deg = degree, TS = trophic similarity, StabDif = stability difference, ADvalue = Anderson-Darling p value.}
\begin{tabular}{l c c c c c c c}
\tophline

\textbf{Species} & \textbf{meanIS} & \textbf{TL} & \textbf{Deg} & \textbf{TS} & \textbf{Habitat} & \textbf{StabDif} & \textbf{ADvalue}\\
\middlehline
Orcinus orca & 0.00163 & 5.03 & 26 & 0.037 & Benthopelagic & 0.00040 & 1.38e-51 \\
\middlehline
Macrourus holotrachys & 0.00074 & 4.70 & 85 & 0.112 & Benthopelagic & 0.00027 & 2.75e-18 \\
\middlehline
Abyssorchomene nodimanus & 0.00022 & 4.21 & 137 & 0.130 & Benthopelagic & 0.00024 & 1.02e-15 \\
\middlehline
Dissostichus mawsoni & 0.00069 & 4.12 & 87 & 0.126 & Pelagic & 0.00018 & 1.06e-10 \\
\middlehline
Pagetopsis macropterus & 0.00063 & 4.64 & 76 & 0.113 & Demersal & -0.00017 & 1.66e- 9 \\
\middlehline
Macrourus whitsoni & 0.00064 & 4.55 & 92 & 0.124 & Benthopelagic & 0.00012 & 4.97e- 8 \\
\middlehline
Hydrurga leptonyx & 0.00092 & 4.72 & 67 & 0.094 & Land-based & 0.00014 & 5.59e- 6 \\
\middlehline
Mesonychoteuthis hamiltoni & 0.00161 & 4.41 & 29 & 0.028 & Pelagic & 0.00013 & 6.53e- 6 \\
\middlehline
Notothenia marmorata & 0.00073 & 4.09 & 44 & 0.091 & Demersal & 0.00014 & 2.10e- 5 \\
\middlehline
Champsocephalus gunnari & 0.00068 & 3.72 & 46 & 0.086 & Pelagic & 0.00010 & 7.97e- 5 \\
\middlehline
Arctocephalus gazella & 0.00083 & 4.67 & 61 & 0.093 & Land-based & 0.00011 & 9.52e- 5 \\
\middlehline
Ommatophoca rossii & 0.00093 & 4.87 & 56 & 0.080 & Land-based & 0.00009 & 1.20e- 3 \\
\middlehline
Notothenia coriiceps & 0.00044 & 4.27 & 130 & 0.126 & Demersal & 0.00012 & 1.33e- 3 \\
